\textbf{How to measure maximally precise radial velocities from a set of spectra:\\ The time-independent spectrum case}

### 2019-06-06 paper/paper.tex
letter
title
abstract, baby

% This file is part of the EPRV project.
% Copyright 2017, 2018 the authors.

@@

\title{\textbf{Data analysis for maximum-precision radial-velocity mesurement}}

@@

\begin{abstract}\noindent
% context
Extremely precise stellar radial velocity (\RV) measurements are a
foundational tool in the field of exoplanets, with hundreds of planets
discovered by the Doppler method to date and many high-resolution
spectrographs around the world dedicated to search and
characterization.
Current observations are delivering \RV\ measurements at $\ms$
precision, while it is thought that some hardware systems ought to be
delivering data capable of even higher precisions.
% aims
The question arises: Is there any precision loss in current
data-analysis methods?
That is, what data-analysis methods deliver \RV\ measurements that
saturate the information-theoretic bound on precision---the
Cram\'er--Rao bound (\CRLB)---given an extracted spectrum (and
associated uncertainties) from a spectrograph?
% methods
We consider cross-correlation with a binary mask, a smoothed (or
Gaussian-fit) cross-correlation with a binary mask, and
cross-correlation with a stellar template.
We consider both theoretically generated and data-driven templates,
and look at some methods for data-driven telluric removal.
% results
As expected, we find that cross-correlation with a correct stellar
template produces \RV\ measurements that saturate the \CRLB.
We find that pure cross-correlation with a binary mask under-performs,
but the standard practice of post-processing the binary-mask result
with a Gaussian fit does saturate the \CRLB, when the binary mask is
``correct'' in a certain sense.
As the template or mask becomes incorrect, \RV\ precision suffers.
Most importantly, we find that using the data themselves to generate
the stellar template does not cost any significant information; that
is, a data-driven template will be correct enough to saturate the
\CRLB, provided that there are multiple epochs of observations and the
star does not vary substantially between epochs.
As a side note, we find that tiny telluric features can affect \RV\ precision
substantially, but that they can be modeled and removed by a data-driven
method, provided that there is sufficient barycentric velocity variation.
\end{abstract}

>>>>

% This file is part of the EPRV project.
% Copyright 2017, 2018, 2019 the authors.

@@

\title{\textbf{How to measure maximally precise radial velocities from a set of spectra:\\
               The time-independent spectrum case}}

@@

\begin{abstract}\noindent
% context
Extremely precise stellar radial velocity (\RV) measurements are a
foundational tool in the field of exoplanets, with hundreds of planets
discovered by the Doppler method to date and many high-resolution
spectrographs around the world dedicated to search and
characterization.
Current observations are delivering \RV\ measurements at $\ms$
precision
It appears that many hardware systems ought to be---on the basis of
information-theoretic considerations---capable of delivering even
higher precisions.
% aims
Here we ask: What data-analysis methods deliver \RV\ measurements that
saturate the
Cram\'er--Rao bound (\CRLB) on precision, given an extracted spectrum (and
associated uncertainties) from a spectrograph?
We presume that the stellar spectrum does not vary with time (this is
incorrect in detail).
% methods
We consider a smoothed (or
Gaussian-fit) cross-correlation with a binary mask,
cross-correlation with a stellar template,
and maximum-likelihood methods.
We draw out the relationships and equivalencies across these methods.
We consider both theoretically generated and data-driven templates.
% results
We find (as expected) that cross-correlation with a correct stellar
template produces \RV\ measurements that saturate the \CRLB.
We also find that 
the standard practice of fitting a Gaussian to the cross-correlation
with a binary mask saturates the \CRLB, so long as the binary mask is
``correct'' in a certain sense.
As the template or mask becomes incorrect, \RV\ precision suffers.
Un-corrected telluric absorption lines can also harm \RV\ precision.
Most importantly, we find that using the data themselves to generate
the stellar and telluric templates does not cost any significant
information.
That is, data-driven stellar and telluric templates can be correct
enough to saturate the \CRLB, provided that there are multiple epochs
of observations and the star does not vary substantially between
epochs.
If the star does vary substantially, a lot more needs to be known
before the relevant \CRLB\ even can be computed.
\end{abstract}
